Theories of deliberated choice

### 2022-11-17 simple.tex
Self-suff using decisive paths (and such def)

While $I$ and $\allargs$ are fundamental sets whose elements are considered elementary, a few notations and concepts are required to define ${\allleadsto}$.
Throughout the article, $\N$ includes zero. %and $\N^* = \N \setminus \set{0}$. 
Given $j, l \in \N$, the notation $\intvl{j, l}$ represents $\set{k \in \N \suchthat j ≤ k ≤ l}$ (thus $\forall j \in \N: \intvl{j, 0} = \emptyset$).
Define $\allhist = \bigcup_{k \in \N} \allargs^{\intvl{1, k}}$ as the finite sequences of arguments, including the empty sequence. 
A generic element of $\allhist$ (a sequence of arguments) is denoted by $\hist$ and called an argumentative path, or simply a path. Its length is denoted by $\card{\hist} \in \N$.
Thus, given an argumentative path $\hist \in \allhist$, $\forall k \in \intvl{1, \card{\hist}}: \hist(k) = \hist_k \in \allargs$, and $\hist = \emptyset ⇔ \card{\hist} = 0$.

@@

An argumentative theory claims that whenever some of its arguments fail to convince $i$ of choosing $\pga$, it can argue something else.
It is said to be truthful iff this claim is correct.
More precisely, its claims are about paths of the sort $\hist = (\arga, …, \ar_2, …, \ar_3, …)$ where $\ar_2 \in \dgi(\arga)$, $\ar_3 \in \dgi(\ar_2)$, …
The following definition makes this precise by mandating that whenever increasing subsequences of $\hist$, involving arguments $\ar_j$ proposed by the theory, are not convincing, then the theory is able to produce yet another argument in defense of its claim.%
\footnote{Intuition.
If $a_1$, used by the theory, is not decisive, then the theory has an answer, namely, $a_2$, which works at least temporarily.
If $a_1$ used by th is not dec, and $a_2$ used by th is not dec *in the same sequence*, then $a_3$ needs to defend *in the same sequence*.
$(…, a_1)$ NO and $(…, a_1, …, a_2)$ NO then $a_3$ defends it.
$\alpha$ of the form $(\alpha^1, \alpha^2)$ where $a_1$ in end of $α^1$ NO and $a_2$ in end of $α^2$ NO.}
\begin{definition}[Truthfulness]
	\label{def:truth}
	An argumentative theory $(\pga, \arga, \dg)$ for the exposure protocol $(I, \allargs, {\leadsto})$ is \emph{truthful} iff 
	$\forall i \in I, \hist \in \allhist, n \in \N, k \in \intvl{1, \card \hist}^n, \ar \in \allargs^n$:
	\begin{equation}
		\left.\begin{aligned}
			&\ar_1 = \arga \\
			&\forall 1 ≤ j < n: \ar_{j + 1} = \dg(\ar_j, α_{\intvl{1, k_j}}) \in \histbend{k_{j + 1}} \\
			&\forall 1 ≤ j ≤ n: \hist_{\intvl{1, k_j}} \nileadsto \pga
		\end{aligned}\right\}
		⇒ \dg(\ar_n, \hist) ≠ \emptyset.
	\end{equation}
\end{definition}
\begin{axiom}[Answerability]
	\label{ax:ans}
	$\forall i \in I, \hist \in \allhist, \ar \in \gargsi$:
	\begin{equation}
		\left.\begin{aligned}
			&\hist \nileadsto \pga \\
			&(\hist, \ar) \ileadsto \pga \\
			&\forall \ar' \in \allargs: (\hist, \ar, \ar') \ileadsto \pga
		\end{aligned}\right\}
		⇒ \ar \ileadstost \pga.
	\end{equation}
\end{axiom}
\footnote{$a_1, a_2, a_3$ NOT and $a_1, a_2, a_3, a_4$ YES and $a_1', a_2, a_3, a_4$ NO imply that $a_1, a_2, a_3, a_4, a'_1$ NO.
Generalizing: $\alpha$ NO and $\alpha, a_4$ YES and $\alpha', a_4$ NO imply that $\exists a' \in \alpha' \suchthat \alpha, a_4, a'$ NO.
Let’s try with this too powerful axiom, but then we should weaken it, as it could be that $a_1$ AND $a_4$ TOGETHER are decisive, whence the axiom fails. We should have then some argument $a_1 + a_4$ to replace them, and…?
In summary, we want to avoid that $a$ is not decisive but we can’t observe it when starting with some $α$ even though $α$ has also no decisive arguments: if $α$ has no decisive argument, then it must be possible to counter it.

@@

\begin{axiom}[Self-sufficiency of positive arguments, version for $\ar$]
	$\forall \ar_1, \ar_2 \in \gargsi, \ar \in \allargs, \hist \in \allhist$:
	\begin{equation}
		\left.\begin{aligned}
			&(\hist, \ar_1, \ar) \nileadsto \pga \\
			&(\hist, \ar_2, \ar) \nileadsto \pga
		\end{aligned}\right\}
		⇒ (\hist, \ar_1, \ar_2, \ar) \nileadsto \pga,
	\end{equation}
	and, $\forall \ar_1, \ar_2 \in \gargsi, \hist \in \allhist$:
	\begin{equation}
		\left.\begin{aligned}
			&(\hist, \ar_1) \nileadsto \pga \\
			&(\hist, \ar_2) \nileadsto \pga
		\end{aligned}\right\}
		⇒ (\hist, \ar_1, \ar_2) \nileadsto \pga.
	\end{equation}
\end{axiom}

>>>>

While $I$ and $\allargs$ are fundamental sets whose elements are considered elementary, a few notations and concepts are required to define ${\allleadsto}$.
Throughout the article, $\N$ includes zero. %and $\N^* = \N \setminus \set{0}$. 
Given $j, l \in \N$, the notation $\intvl{j, l}$ represents $\set{k \in \N \suchthat j ≤ k ≤ l}$ (thus $\forall j \in \N: \intvl{j, 0} = \emptyset$).
Define $\allhist = \bigcup_{k \in \N} \allargs^{\intvl{1, k}}$ as the finite sequences of arguments, including the empty sequence. 
A generic element of $\allhist$ (a sequence of arguments) is denoted by $\hist$ and called an argumentative path, or simply a path. Its length is denoted by $\card{\hist} \in \N$ and its elements, $\hist(k)$ or $\hist_k$.
Thus, given an argumentative path $\hist \in \allhist$, $\forall k \in \intvl{1, \card{\hist}}: \hist(k) = \hist_k \in \allargs$, and $\hist = \emptyset ⇔ \card{\hist} = 0$.

@@

An argumentative theory claims that whenever some of its arguments fail to convince $i$ of choosing $\pga$, it can argue something else.
It is said to be truthful iff this claim is correct.
More precisely, its claims are about paths of the sort $\hist = (\arga, …, \ar_2, …, \ar_3, …)$ where $\ar_2 \in \dgi(\arga)$, $\ar_3 \in \dgi(\ar_2)$, …
The following definition makes this precise by mandating that whenever increasing subsequences of $\hist$, involving arguments $\ar_j$ proposed by the theory, are not convincing, then the theory is able to produce yet another argument in defense of its claim.%
\footnote{Intuition.
If $a_1$, used by the theory, is not decisive, then the theory has an answer, namely, $a_2$, which works at least temporarily.
If $a_1$ used by th is not dec, and $a_2$ used by th is not dec *in the same sequence*, then $a_3$ needs to defend *in the same sequence*.
$(…, a_1)$ NO and $(…, a_1, …, a_2)$ NO then $a_3$ defends it.
$\alpha$ of the form $(\alpha^1, \alpha^2)$ where $a_1$ in end of $α^1$ NO and $a_2$ in end of $α^2$ NO.}
\begin{definition}[Truthfulness]
	\label{def:truth}
	An argumentative theory $(\pga, \arga, \dg)$ for the exposure protocol $(I, \allargs, {\leadsto})$ is \emph{truthful} iff 
	$\forall i \in I, \hist \in \allhist, n \in \N, k \in \intvl{1, \card \hist}^n, \ar \in \allargs^n$:
	\begin{equation}
		\left.\begin{aligned}
			&\ar_1 = \arga \\
			&\forall 1 ≤ j < n: \ar_{j + 1} = \dg(\ar_j, α_{\intvl{1, k_j}}) \in \histbend{k_{j + 1}} \\
			&\forall 1 ≤ j ≤ n: \hist_{\intvl{1, k_j}} \nileadsto \pga
		\end{aligned}\right\}
		⇒ \dg(\ar_n, \hist) ≠ \emptyset.
	\end{equation}
\end{definition}
Given the paths $\alpha, \alpha'$, the sequence $(\alpha, \alpha')$ denotes a path constituted of the arguments in $\alpha$ followed by the arguments of $\alpha'$.
\begin{axiom}[Answerability]
	\label{ax:ans}
	$\forall i \in I, \hist \in \allhist, \ar \in \gargsi$:
	\begin{equation}
		\left.\begin{aligned}
			&\hist \nileadsto \pga \\
			&(\hist, \ar) \ileadsto \pga \\
			&\forall \ar' \in \allargs: (\hist, \ar, \ar') \ileadsto \pga
		\end{aligned}\right\}
		⇒ \ar \ileadstost \pga.
	\end{equation}
\end{axiom}
\footnote{$a_1, a_2, a_3$ NOT and $a_1, a_2, a_3, a_4$ YES and $a_1', a_2, a_3, a_4$ NO imply that $a_1, a_2, a_3, a_4, a'_1$ NO.
Generalizing: $\alpha$ NO and $\alpha, a_4$ YES and $\alpha', a_4$ NO imply that $\exists a' \in \alpha' \suchthat \alpha, a_4, a'$ NO.
Let’s try with this too powerful axiom, but then we should weaken it, as it could be that $a_1$ AND $a_4$ TOGETHER are decisive, whence the axiom fails. We should have then some argument $a_1 + a_4$ to replace them, and…?
In summary, we want to avoid that $a$ is not decisive but we can’t observe it when starting with some $α$ even though $α$ has also no decisive arguments: if $α$ has no decisive argument, then it must be possible to counter it.

@@

A path can be shortened to its first $k \in \intvl{0, \card{\hist}}$ elements, which is denoted by $\restr{\hist}{\intvl{1, k}}$, and “shortening” a path beyond its length yields the same path again, thus given $k \in \N$, $\restr{\hist}{\intvl{1, k}} = \restr{\hist}{\intvl{1, \min\set{\card{\hist}, k}}}$.
Define a path as being decisive for $p$, $\hist \ileadstost p$, iff $\hist \ileadsto p \land \forall \ar \in \allargs: (\hist, \ar) \ileadsto p$, or equivalently, iff $\forall \hist' \suchthat \card{\hist} ≤ \card{\hist'} ≤ \card{\hist} + 1 \land \restr{\hist'}{\intvl{1, \card{\hist}}} = \hist: \hist' \ileadsto p$. Thus, $\hist \nileadstost p ⇔ \exists \ar \in \allargs \suchthat (\hist, \ar) \nileadsto p$.

\begin{axiom}[Self-sufficiency of positive arguments]
	$\forall \ar_1, \ar_2 \in \gargsi, \ar \in \allargs, \hist \in \allhist$:
	\begin{equation}
		\left.\begin{aligned}
			&(\hist, \ar_1) \nileadstost \pga \\
			&(\hist, \ar_2) \nileadstost \pga
		\end{aligned}\right\}
		⇒ (\hist, \ar_1, \ar_2) \nileadstost \pga.
	\end{equation}
\end{axiom}
